### 2017-11-20 linearity-2018/introduction.tex
Finished the introduction, and moving on to:
The first section on adjoint models, and then the systems.

Perhaps the most elegant model of intuitionistic linear logic is
Benton's linear/non-linear (LNL) models \cite{Benton:1994}. On the
semantic side, LNL models are very compact and intuitive, but on the
syntactic side, they correspond to a logic and type theory, called LNL
logic, that allows one to reason with or without weakening and
contraction without the need to annotate every formula of the logic
with the of-course!  modality.  This is possible, because formulas
like $A \to B$ are primitives of the logic, and not encoded using the
of-course! modality.  Therefore, we ask the question, ``can a similar
elegant, intuitive, and flexible model and corresponding logic be
defined for the Lambek Calculus?''

@@

Linguists have wondered if a modality for exchange can be added to the
Lambek Calculus similarly to how the of-course! modality adds
weakening and contraction to linear logic.  de Paiva and
Eades~\cite{?} show that this is possible by adding a modality $\kappa
A$ with the property that $(\kappa A \otimes B) \multimap (B \otimes
\kappa A)$.  The adjoint model introduced above induces the modality
$eA = F(G\,A)$ with the property $eA \otimes eB \multimap eB \otimes
eA$.  Thus providing a second solution to the problem, but we go
further, and show such a modality can be given for associativity. It
arises as the induced comonad of the adjoint model where $\cat{M}_1$
is a monoidal category, and $\cat{M}_2$ is a vanilla magmoidal
category with a unit.  Then we will obtain the property $(aA \otimes
aB) \otimes aC) \multimap aA \otimes (aB \otimes aC)$.

Linguists as well as computer scientists have found many applications
of substructural logics that contain more than one tensor product
satisfying different structural rules.  For example, a commutative
tensor product and a non-commutative tensor product within the same
logic.  An example of such a logic is the logic of bunched
implications, but there sequents $\Gamma \vdash A$ depend on contexts
$\Gamma$ that are trees with multiple different branching nodes
instead of lists.  The branching nodes are necessary to accommodate
multiple different tensor products with different structural
properties.  We show how adjoint models can be used to build similar
logics, but without the need for multiple different branching nodes.
Furthermore, adjoints compose, and this property can be exploited to
compose modalities.

%% Security, concurrency, and PL motivation

>>>>

Perhaps the most elegant model of intuitionistic linear logic is
Benton's linear/non-linear (LNL) models \cite{Benton:1994}. On the
semantic side, LNL models are very compact and intuitive, but on the
syntactic side, they correspond to a logic and type theory, called LNL
logic, that allows one to reason with or without weakening and
contraction without the need to annotate every formula of the logic
with the of-course!  modality.  This is possible because formulas like
$A \to B$ are primitives of the logic, and not encoded using the
of-course! modality.  Therefore, we ask the question, ``can a similar
elegant, intuitive, and flexible model and corresponding logic be
defined for the Lambek Calculus?''

@@

Linguists have wondered if a modality for exchange can be added to the
Lambek Calculus similarly to how the of-course! modality adds
weakening and contraction to linear logic.  de Paiva and
Eades~\cite{?} show that this is possible by adding a modality $\kappa
A$ with the property that $(\kappa A \otimes B) \multimap (B \otimes
\kappa A)$.  The adjoint model introduced above induces the modality
$eA = F(G\,A)$ with the property $eA \otimes eB \multimap eB \otimes
eA$.  Thus providing a second solution to the problem, but we go
further, and show such a modality can be given for associativity. It
arises as the induced comonad of the adjoint model where $\cat{M}_1$
is a monoidal category, and $\cat{M}_2$ is a vanilla magmoidal
category with a unit.  Then we will obtain the isomorphism $(aA
\otimes aB) \otimes aC) \multimap aA \otimes (aB \otimes aC)$.

Linguists as well as computer scientists have found many applications
of substructural logics that contain more than one tensor product
satisfying different structural rules.  For example, a commutative
tensor product and a non-commutative tensor product within the same
logic.  An example of such a logic is the logic of bunched
implications, but there sequents $\Gamma \vdash A$ depend on contexts
$\Gamma$ that are trees with multiple different branching nodes
instead of lists.  The branching nodes are necessary to accommodate
multiple different tensor products with different structural
properties.  We introduce, but leave the details to future work, the
idea of using adjoint models to build similar logics, but without the
need for multiple different branching nodes.  Furthermore, adjoints
compose, and this property can be exploited to compose modalities.

We introduce the idea above of having a modality for exchange and one
for associativity, but what about weakening and contraction?  Indeed
it is possible to give modalities for just weakening, $wA$, and just
contraction, $cA$, using adjoint models.  Now that we have each
structural rule isolated into their own modality is it possible to put
them together to form new modalities that combine structural rules?
The answer to this question has already been show to be positive, at
least for weakening and contraction, but we extend this line of work
to include exchange and associativity as well.
Jacobs~\cite{JACOBS199473} used monads and distributive laws to relate
modalities, but Melli{\'e}s~\cite{Mellies:2004} shows how to use
adjunctions to accomplish the same, but in a more intuitive and
natural way.  Thus, this part of our work is a natural extension of
Melli{\'e}s'.  At this point we are starting to see that this is a
very general framework.

This work can be used to define a general core logic that is capable
of encoding several different substructural logics.  Some examples are
intuitionistic linear logic, affine logic, contraction logic, and the
(non-associative) Lambek Calculus.  This does have some similarities
to Licata et al.~\cite{licata2017fibrational}, but there they embrace
the tree based structures with multiple different branching node for
contexts from the logic of bunched implications.  However, our work is
trying to understand if this is truly necessary, and if adjunctions
provide a means of, at the very least, hiding these multiple branching
nodes.  Thus, adjunctions may lead to a simplification.
